Busch, Khakhalin. Intrinsic temporal tuning

### 2019-02-02 reviews and responses.md

Reviews and Authors' Responses for (Busch Khakhalin 2018)
======================

This is a collection of reviews and feedbacks, both official and unofficial, for different working copies of this paper. Reviews are split into paragraphs, and Author's responses and comments are shown after each paragraph. Reviews, and responses to these revivews, are given in reverese chronological order.

# Reviews from J Neurosci, 2018 Nov 01 (rejection letter)

These reviews were received on Nov 1st 2018 from J Neurosci, as this paper was rejected by the editor. We decided to include these reviews to our next submission, and to this repository, and provide our responses to each of the critiques.

@@

> A reasonable attempt was made explain the conditioning changes observed at a more mechanistic level, but at least for the dynamic clamp experiments the authors were unable to account for their observations based on traditional mechanisms for regulation of intrinsic excitability. Overall the paper is fascinating and introduces a potentially interesting new way of thinking about plasticity, but doesn't quite provide enough of a mechanistic underpinning nor of a functional context for the model. It comes across as a slightly contrived set of systematic observations. The paper has a few minor technical issues, but suffers mainly from failing to systematically test any hypothesis in a satisfying and reductionist way. The paper would likely be better suited either for a more specialized cellular neurophysiology journal or, if the narrative can be tightened up to clearly convey its novelty, for a short format high-impact report.

&nbsp;

We are grateful for the kind words, and agree with the assessment that our paper is largely phenomenological. Unfortunatelly, in a small teaching-oriented college, we simply cannot run most experiments needed to probe the mechanisms of intrinsic temporal tuning. We cannot perform in-vivo experiments with sensory stimulation; and cannot measure or change gene expression in isolated tectal cells. Even a detailed electrophysiological study of isolated ionic currents would have taken extra 2-3 years, so we decided to publish this phenomenological report, in a hope that the very phenomenon we describe is interesting enough to be followed up by other labs, potentially in different models.

In this, updated version of the paper, we try to be more explicit about possible next steps that may be taken to study the mechanisms behind the exciting phenomenon we discovered.

@@

> 3.	Page 3 line 18, "when tadpoles were exposed to" -> "in tadpoles that had been exposed to"

Corrected.

> 4.	The statistical tests applied throughout the paper are unclear and need more explanation. The number of degrees of freedom (e.g, 667) seems far too high, suggesting that multiple recordings from the same cell under different conditions are being treated as independent N. However recordings from the same cell under different temporal or amplitude conditions are far from independent measurements. It seems that a two-way or even three-way ANOVA might be the best approach for examining the variables of conditioning stimulus and response property.

The number of degrees of freedom in question comes from a full linear model, and does not mean that we treated different conditions as independent. Typically, to compare the numer of spikes in two conditions we used a 3-way ANCOVA (injection amplitude and duration as ordinal variables; group id as a nominal variable) with interactions, and repeated measures (where cell id included as a nominal variable). For example, for a comparison of "Control" group to "Sound" group we used the following R code (see `dynamic_spikes_plot.R` for the full code):

@@

```R
ds = subset(d,is.element(Group,c("Control","Sound")))
summary(aov(data=ds,Spikes~Shape*Amp + Shape*Group + Amp*Group + Cell*Shape*Amp))
```

In addition to a full description of our analysis approach, as given in the "Methods", we now included a second, shoter description of it in the "Results" section, to make the paper more readable. We also improved the description in the "Methods" section; and added several extra referrals to the "Methods" section from the "Results" section, to remind the reader that extra information is available there. In particular, in the sentence that reports the first F-value for this study, we now added the following commentary:

@@

Fixed.

> 7.	The narrative logic on page 9 lines 11-22 is confusing. How does homeostatic plasticity, if it truly is at play here, "increase their ability to respond to unusual patterns of synaptic activation." If anything wouldn't homeostasis decrease dynamic range by pushing all cells toward a central set point?

We agree that the word "homeostatic" was confusing in this context. We now simplified the sentence:

"This means that individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, decreasing their responses to common input patters, and enhancing responses to unusual patterns of synaptic activation."

@@

> 8.	Page 12 the section on position within the tectum feels like it was just tacked onto the end of the paper. This should be integrated earlier in the paper perhaps to go with figure 1A.

We placed the decription of this effect at the very end of the paper, as we prefer to downplay it a bit, as it was described previously (Khakhalin 2012; Hamodi 2012; Ciarleglio 2015), and we don't want to distract the reader from our other findings. At the same time, we find it important to mention that we observed a strong correlation with position: first, because it provides an independent replication of previously observed patterns, and second, as the dependency of cellular properties on position necessitated an adjustment for position during analysis (see "Methods"), which needed to be justified. The implication for the principle of parameter degeneracy in development (Prinz 2004) is also interesting, and we wanted to point at it, even though within this paper we don't have enough material to fully explore this theme.

@@

> The paper is, in my opinion, hard to read because the authors are often not clear about what they mean, refer to previous parts of the paper assuming the reader has a perfect memory or is willing to turn back and forth across pages, and use non-traditional terminology. In addition, the statistically analysis appears to be pooling many data points from non-independent cells, rendering many of the conclusions invalid. Further, the experiments intended to measure intrinsic properties (in Figures 1 and 2) appear to have been performed without synaptic blockers, so the the adaptation that they measure could be in the circuit (inhibitory recruitment) instead of in the intrinsic properties.

> There is indeed some merit in the goals of the study (to see if adaptation changes with experience in addition to conductance/response gains), but it has so many technical problems at present that it would need to be radically different (perhaps with experiments and analysis re-done) to perhaps be of interest to the broad readership of J Neuroscience.

&nbsp;

On excessive cross-references between different parts of the paper: thank you for pointing it out! We now tried to link different parts of the paper more tightly together, by gently reminding the reader about previously made statements every time we make a reference.

On statistical analysis: we do not pool data points from non-independent cells, but used a repeated measures analysis, verified by mixed model analysis (see Methods, as well as corrections described above, in our responses to Reviewer 1).

On synaptic blockers: it is true that we measured intrinsic properties without synaptic blockers, but this is considered a standard practice in this preparation, as most earlier studies measured IV curves without synaptic blockers (Azenman .. Cline 2002; Aizenman .. Cline 2003; James .. Aizenman 2015; Ciarleglio Khakhalin .. Aizenman 2015). What makes this practice possible is a relatively weak recruitment of polysynaptic activity after activation of a single tectal cell (Bell .. Aizenman 2011).

@@

> Page 8, lines 2-12. "Average firing rate" is a more common term than "spikiness".

Please see our comments above. The term "average firing rate" is typically used to refer to actual average firing rates, either long-term or instaneous, in situations when relatively realistic neural activity is possible, such as in-vivo recordings, or during spontaneous activity. We however looked at a rather artificial excitability index, and we would prefer to stress that by naming this index, and defining it within the study.

> How about "gain" for amplitude tuning? I see that the authors are using the parameter a to indicate the non-linearity in the tuning curve. For "temporal tuning", why not use "adaptation index" or something like that, that could be plotted in Figure 1F? The "temporal tuning" is a measure of non-linearity, but it is not very intuitive. (It is not "incorrect" as the authors have it now, but I think it would be hard for the interdisciplinary readership of J Neuroscience to get on board easily.)

We would really prefer to retain terms "amplitude tuning" and "temporal tuning", as they better represent the main message of our paper, as we are trying to think of these values not in terms of abstract properties each neuron has, but as parameters that can tune a neuron to a certain type of synaptic inputs, in terms of their temporal properties, and their amplitude. We however now included words "gain" and "adaptation index" in paragraphs that introduce values of amplitude and temporal tuning, respectively.

@@

> Figure 2: Units needed for Y axis on A, B (units of a are spikes per conductance squared and units of amplitude tuning are spikes/conductance)

Both tuning parameters are dimensionless, as both conductance and amplitude were treated as rank-transformed, ordinal values (see methods).

@@

> Figure 2B and Figure 3A, 3D don't show any post-hoc comparisons among the groups. Which groups differ? I see it in the text but it is hard to take it home.

We added asterisks to mark significance in figures.

> Figure 3E and G: how can we reconcile these accounts? In each of the groups (3G), there is at best a very weak relationship between temporal tuning and synaptic duration. Somehow, if you take the means of all of these groups, you can fit a line through them (Figure 3E). Why is that the right thing to do? Aren't the field of individual points (partially-transparent background) and the data plotted in 3G the relative quantities?

We now added a paragraph describing this situation in detail, and reference a textbook name for this type of situation, when between-group and within-group analyses yield seemily opposite results (Simpson's paradox).

@@

"As expected, cells with strong monosynaptic inputs had shorter synaptic currents, while polysynaptic activity made synaptic currents longer"

---------- TODO ----------- : Add a paragraph to the paper, explaining that synaptic duration is necessary, as it is what allows us to compare apples to apples, as we fit the synaptic currents exactly with a waveform we used to assess temporal tuning in dynamic clamp experiments. Then reiterate, and cite this paragraph here.

> Figure 3G, page 10 lines 24-40. Creation of "super groups" is totally bogus. This part should be dropped. I see no strong significant relationship between temporal tuning and synaptic duration, and am not convinced synaptic duration is a meaningful quantity.

Removed. We attempted a different, more convincing analysis of the interaction between synaptic and intrinsic properties across groups, but when split by group, our analysis lacks power, so we decided to follow the advice and remove this paragraph.

@@

# Reviews from eLife, 2018 Aug 17 (rejection letter)

> ...While both BRE members found much to like about the manuscript, they found it dense going. But they both felt this paper belongs in a more specialized journal. A comment from one of them:

> The authors pose two questions in the introduction: 1. Is intrinsic plasticity limited to changes in excitability? 2. Whether intrinsic plasticity is a response to synaptic activation? There are data in other systems that make these questions "straw men". We know from the STG that intrinsic plasticity is not limited to change in spikiness alone and that there are several other parameters such as bursts, rebounds, spike widths etc which can and do get altered in the short and long term. There are also plenty of examples where one ion channel expression is co-regulated with another type of ion channel, or by slow action of growth factors, hormones and other neuromodulators. These are examples of intrinsic plasticity not governed by fast synaptic transmission alone.

This is of course true, but while many types of intricate effects of intrinsic plasticity are described in small networks, and especially in oscillatory networks, such as the STG, it is less common to consider these non-trivial effects in large feed-forward excitatory networks, such as networks in the cortex, or superior colliculus. It is natural for researchers to limit the complexity of their conceptual and computational models, and it is probably fair to say that as the number of neurons involved in a model increases, researchers often shift their attention from intrinsic mechanisms to connectomic approaches, approximating neural networks as a collection of nodes, linked with synapses of different strengths (e.g: Cline 2008; Kirby 2013). Even realistic biophysical models that represent neurons as multi compartment systems with active dendritic currents often don't assume that these active currents may be modulated in response to activation (Moldwin 2018). When machine learning approaches are used as a model for brain development, intrinsic excitability is reflected in these models, as a change in activation function of each artificial neuron (Marblestone 2016). This type of modeling implicitly assumes, however, that outside of obviously oscillatory networks (spinal cord, thalamus, etc.), intrisic excitability in biological networks is largely confined to overall changes in neuronal "responsiveness".

### 2019-02-02 main.tex
tex: figure captions, responses to reviewers

\begin{figure*}
\includegraphics[width=\linewidth]{fig1.png}
\caption{
Overview of experimental design, and summary of dynamic clamp results. (\textbf{A}). Positions of tectal neurons that were recorded. (\textbf{B}). Sample data from a dynamic clamp experiment. Bottom row: the dynamics of conductances G(t) of four different durations simulated by the dynamic clamp system. Middle row: the currents I(t) dynamically injected into a cell based on conductances of 4 different durations and 3 different amplitudes. Top row: resulting voltage traces V(t) that were recorded and analyzed. (\textbf{C}). A schematic of visual conditioning, in “Flash” (left) and “Looming” (right) groups. (\textbf{D}). The number of spikes produced by all neurons in all experiments, split by input peak conductance, and plotted against conductance duration. Black lines show respective averages. (\textbf{E}). A summary of data from (d), presented as averages and 95\% confidence intervals.}
\end{figure*}

@@

We then looked at how an increase in transmembrane conductance translated into increased spike output of neurons, to map their amplitude tuning, or amplitude transfer function. Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve, and did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming transitions respectively). The difference in amplitude tuning between these two types of stimulation was not significant (F(1,641)=3.5; p=0.06).

@@

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig2.png}
\caption{
Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to a plateau-shaped curve (purple, yellow). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups. (\textbf{C}). Temporal tuning and average spikiness of neurons in different experimental groups; in each plot all neurons across all groups are shown in gray, while neurons from one target group are shown in color; means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control, and when they were added to “Flashes”, to form two types of multisensory stimuli.}
\end{figure*}

@@

To visualize and interpret differences in neuronal tuning, we quantified each of the three aspects of intrinsic tuning (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value of "temporal tuning" (Figure 2A). The numerical values of these "tuning coefficients" are not easily interpretable, but they capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see "Methods" for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009; see "Methods" for a description of the linear model we used). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.46 and 0.74 on log-transformed data respectively). The amplitude of late synaptic currents–produced by recurrent network activation (15-145 ms after the stimulus)–did not differ across groups.

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig3.png}
\caption{Changes in synaptic transmission, and co-tuning of synaptic and intrinsic neuronal properties. (\textbf{A}). Amplitudes of early monosynaptic inputs to tectal neurons in all experimental groups (given in log scale, with outliers brought within the axes limits). (\textbf{B}). A sample synaptic recording, showing all traces (green) and an average trace (blue). The black bars show the areas at which early monosynaptic and late polysynaptic currents were measured; the vertical position of each bar represents the respective average current. The second, longer bar does not completely fit within the figure at this scale. (\textbf{C}). Synaptic current duration (vertical axis) was mostly defined by the amplitude of early monosynaptic inputs (horizontal axis). (\textbf{D}). Synaptic current duration was different between experimental groups (see the text). (\textbf{E}). Across experiments, average temporal tuning in each group positively correlated with the average durations of synaptic currents they received. (\textbf{F}). Within experimental groups, temporal tuning of individual neurons negatively correlated with the duration of synaptic currents they received. Axes show within-group deviations of temporal tuning and synaptic current duration from respective averages for each group. (\textbf{G}). Raw tuning and synaptic duration values for each treatment group, with linear regression lines.}
\end{figure*}

Neurons with different contribution of early and late synaptic responses naturally had different synaptic current duration, which we quantified for every cell as a “center of mass” of the current within the first 700 ms after optic chiasm stimulation (see “Methods”). As expected, cells with strong monosynaptic inputs had shorter synaptic currents, while polysynaptic activity made synaptic currents longer (Figure 3C; p=2e-16, r=$-$0.78, n=168). We found that the synaptic current duration was different across treatment groups (Figure 3D; F(5,163)=6.3, p=2e-5). Cells in Flash, Sync, and Async groups all received shorter synaptic inputs than Control cells (Tukey p$<$0.05, mean duration of 267$\pm$36, 253$\pm$39, 268$\pm$24 and 304$\pm$40 ms respectively), indicating that prolonged sensory activation with short, frequent stimuli, reshaped synaptic transmission in the tectum, making it faster, because of selective potentiation of visual inputs from the eye.

@@

In contrast, \textit{within} experimental groups, cells that preferred shorter synaptic inputs tended to receive longer synaptic currents, and vice versa (F(1,145)=4.9, p=0.03). We calculated differences between the properties of each individual cell and the average for the experimental group it belonged to, and showed these "deviations from respective averages" on one plot (Figure 3F). Cells that had shorter synaptic currents, compared to other cells in their group, tended to be selective for longer synaptic currents (adjusted r=$-$0.18, p=0.03, n=151). This means that individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, increasing their ability to respond to unusual patterns of synaptic activation.

%%% --- This section below was commented out in prior submissions

Note that the correlation of intrinsic and synaptic properties had opposite signs between-groups (positive) and within-groups (negative), so if we lumped all cells from all groups together, we would have found no correlation between these two variables (r=$-$0.04, p=0.7, n=151). This is a textbook example of a so called "Simpsons paradox", wherein a pattern holds within subgroups, but disappears or is reversed on a full set because of pronounced differences between groups.

To better understand the interplay between the stimulation-induced retuning of the entire tectal network, and tuning of individual cells within each network, we looked at the interaction of intrinsic and synaptic temporal properties separately for each treatment group (Figure 3G). The co-tuning seemed present in Control, Looming, and Sound groups (r=$-$0.11, $-$0.36, and $-$0.47), but was absent in groups that experienced flashes, and so were subjected to stronger retuning (r=0.27, $-$0.01, and 0.01 for Flash, Sync, and Async groups respectively). If three slow, and three fast (Flash-like) groups were combined into two larger "supergroups", the Control-Looming-Sound neurons exhibited a homeostatic negative correlation between synaptic duration and intrinsic temporal tuning (r=$-$0.29, p=0.01, n=70), while the Flash-Sync-Async neurons lacked cells preferring slow stimuli, and so lacked the co-tuning (r=0.08, p=0.50, n=80). This suggests that when fast visual stimulation made the tectal network selective to short activation patterns, it happened not because all neurons uniformly changed their properties, but because a particular subset of neurons, namely those that receive strong direct inputs from the retina and weak recurrent inputs, become less excitable. Normally these neurons experience fast synaptic inputs, but are most responsive to slow conductances; in overstimulated networks however they became selective to fast conductances.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.2in]{fig4.png}
\caption{Electrophysiological properties of individual neurons, compared to their spiking in current and dynamic clamp experiments. (\textbf{A}). Sample curves for a voltage step experiment; black bars show the areas used to average Na (bottom) and transient K (top) currents (see Methods). (\textbf{B}). Processing of ionic currents data, with IV-curves translated into two parameters (threshold potential and peak current) for each ionic conductance. (\textbf{C}). Sample data from a current clamp experiment; spikes are marked with black dots. (\textbf{D}). Estimations of spikiness from current clamp experiments (horizontal axis) and dynamic clamp experiments (vertical axis) correlate. (\textbf{E}). The number of spikes registered in current clump mode: values predicted from a linear model, plotted against actually observed values. The model works reasonably well (61\% of variance explained). (\textbf{F}). Similar comparison for the dynamic clamp experiments: the model has very low predictive value (13\% of variance explained). Here and in E, both values are adjusted for position.}
\end{figure}

>>>>

\begin{figure*}
\includegraphics[width=\linewidth]{fig1.pdf}
\caption{
Overview of experimental design, and summary of dynamic clamp results. (\textbf{A}). Positions of tectal neurons that were recorded. (\textbf{B}). Sample data from a dynamic clamp experiment. Bottom row: the dynamics of conductances G(t) of four different durations simulated by the dynamic clamp system. Middle row: the currents I(t) dynamically injected into a cell based on conductances of 4 different durations and 3 different amplitudes. Top row: resulting voltage traces V(t) that were recorded and analyzed. (\textbf{C}). A schematic of visual conditioning, in “Flash” (left) and “Looming” (right) groups. (\textbf{D}). The number of spikes produced by all neurons in all experiments, split by input peak conductance, and plotted against conductance duration. Black lines show respective averages. (\textbf{E}). A summary of data from (d), presented as averages and 95\% confidence intervals.}
\end{figure*}

@@

We then looked at how an increase in transmembrane conductance translated into increased spike output of neurons, to map their amplitude tuning (also known as amplitude transfer function, or gain). Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve, and did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming transitions respectively). The difference in amplitude tuning between these two types of stimulation was not significant (F(1,641)=3.5; p=0.06).

@@

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig2.pdf}
\caption{
Quantification of changes in temporal tuning in response to sensory experience. (\textbf{A}). An illustration of how “Temporal tuning” and “Amplitude tuning” values were calculated. For the temporal tuning measure, the value of zero corresponds to linear dependency (blue line), positive values - to an accelerating, supralinear curve (red), and negative values - to a plateau-shaped curve (purple, yellow). For amplitude tuning, higher values correspond to faster increase in spiking with increased conductance. (\textbf{B}). Amplitude tuning of neurons across different experimental groups (stars show t-test p<0.05 compared to control). (\textbf{C}). Temporal tuning and average spikiness of neurons in different experimental groups; in each plot all neurons across all groups are shown in gray, while neurons from one target group are shown in color; means are shown as black dots; ellipses represent 95\% normal confidence regions. Two outliers (top right corner) are brought within the axes limits. (\textbf{D}). Same data as in (C), shown as averages for each group, with 95\% confidence intervals. Black arrows show the effects of sound clicks, when they were added to control, and when they were added to “Flashes”, to form two types of multisensory stimuli.}
\end{figure*}

@@

To visualize and interpret differences in neuronal tuning, we quantified each of the three aspects of intrinsic tuning (average spikiness, amplitude tuning, and temporal tuning) with one value per neuron (see Methods). We used the mean number of spikes across all conditions as the measure of "spikiness"; the linear slope of the number of spikes as a function of conductance amplitude as the measure of "amplitude tuning", and the quadratic term of the curvilinear regression for the number of spikes as a function of input duration as the value characterising the "temporal tuning" or adaptation index of each neuron (Figure 2A). The numerical values of these "tuning coefficients" are not easily interpretable, but they capture the character of tuning curves for each neuron (Figure 2A). All three parameters differed across experimental groups: F(5,160)=3.1, p=0.01 for average spikiness (Figure 2C; see "Methods" for model description); F(5,160)=4.8, p=4e-4 for amplitude tuning (Figure 2B); and F(5,160)=3.6, p=4e-3 for temporal tuning (Figure 2C).

@@

To see whether prolonged sensory stimulation affected synaptic inputs received by tectal neurons, we recorded evoked excitatory postsynaptic currents in response to optic chiasm stimulation. We found that the amplitude of the early, monosynaptic component of evoked responses (the average current between 5 and 15 ms after the shock; Figure 3B) differed across experimental groups (Figure 3A; F(5,161)=3.2, p=0.009; see "Methods" for a description of the linear model we used). Both Sync and Async multisensory groups had larger early synaptic currents than the Control group (Tukey p=0.03 and 0.04; Cohen d=0.92 and 0.73 on log-transformed data respectively). The amplitude of late synaptic currents–produced by recurrent network activation (15-145 ms after the stimulus)–did not differ across groups.

\begin{figure*}[!t]
\includegraphics[width=\linewidth]{fig3.pdf}
\caption{Changes in synaptic transmission, and co-tuning of synaptic and intrinsic neuronal properties. (\textbf{A}). Amplitudes of early monosynaptic inputs to tectal neurons in all experimental groups (in log scale, outliers brought within the axes limits, stars show t-test p<0.05 compared to control). (\textbf{B}). A sample synaptic recording, showing all traces (green) and an average trace (blue). The black bars show the areas at which early monosynaptic and late polysynaptic currents were measured; the vertical position of each bar represents the respective average current. The second, longer bar does not completely fit within the figure at this scale. (\textbf{C}). Synaptic current duration (vertical axis) was mostly defined by the amplitude of early monosynaptic inputs (horizontal axis). (\textbf{D}). Synaptic current durations were different between experimental groups (see text; stars show t-test p<0.05 compared to control). (\textbf{E}). Across experiments, average temporal tuning in each group positively correlated with the average durations of synaptic currents they received. (\textbf{F}). Within experimental groups, temporal tuning of individual neurons negatively correlated with the duration of synaptic currents they received. Axes show within-group deviations of temporal tuning and synaptic current duration from respective averages for each group.}
\end{figure*}

Neurons with different contribution of early and late synaptic responses naturally had different synaptic current duration, which we quantified for every cell as a “center of mass” of the current within the first 700 ms after optic chiasm stimulation (see “Methods”). As expected, cells with strong monosynaptic inputs had shorter synaptic currents, while polysynaptic activity made synaptic currents longer (Figure 3C; p=2e-16, r=$-$0.78, n=168). The synaptic current duration was different across treatment groups (Figure 3D; F(5,163)=6.3, p=2e-5); cells in Flash, Sync, and Async groups all received shorter synaptic inputs than Control cells (Tukey p$<$0.05, mean duration of 267$\pm$36, 253$\pm$39, 268$\pm$24 and 304$\pm$40 ms respectively), indicating that prolonged sensory activation with short, frequent stimuli, reshaped synaptic transmission in the tectum, making it faster, because of selective potentiation of visual inputs from the eye.

@@

In contrast, \textit{within} experimental groups, cells that preferred shorter synaptic inputs tended to receive longer synaptic currents, and vice versa (F(1,145)=4.9, p=0.03). We calculated differences between the properties of each individual cell and the average for the experimental group it belonged to, and showed these "deviations from respective averages" on one plot (Figure 3F). Cells that had shorter synaptic currents, compared to other cells in their group, tended to be selective for longer synaptic currents (adjusted r=$-$0.19, p=0.02, n=151). This means that individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, decreasing their responses to common input patters, and enhancing responses to unusual patterns of synaptic activation.

%%% --- The paragraph below was commented out in prior submissions

Note that the correlation of intrinsic and synaptic properties had opposite signs between-groups (positive) and within-groups (negative), so if we lumped all cells from all groups together, we would have found no correlation between these two variables (r=$-$0.07, p=0.4, n=151). This is a textbook example of a so called "Simpsons paradox", wherein a pattern holds within subgroups, but disappears or is reversed on a full set because of pronounced differences between groups.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.2in]{fig4.pdf}
\caption{Electrophysiological properties of individual neurons, compared to their spiking in current and dynamic clamp experiments. (\textbf{A}). Sample curves for a voltage step experiment; black bars show the areas used to average Na (bottom) and transient K (top) currents (see Methods). (\textbf{B}). Processing of ionic currents data, with IV-curves translated into two parameters (threshold potential and peak current) for each ionic conductance. (\textbf{C}). Sample data from a current clamp experiment; spikes are marked with black dots. (\textbf{D}). Estimations of spikiness from current clamp experiments (horizontal axis) and dynamic clamp experiments (vertical axis) correlate. (\textbf{E}). The number of spikes registered in current clump mode: values predicted from a linear model, plotted against actually observed values. The model works reasonably well (61\% of variance explained). (\textbf{F}). Similar comparison for the dynamic clamp experiments: the model has very low predictive value (13\% of variance explained). Here and in E, both values are adjusted for position.}
\end{figure}
